TensorAlloy: an automatic atomistic neural network program for alloys

### 2019-07-17 doc/papers/nn/manuscript.tex
Made some improvements to the revision.

\title{
    TensorAlloy: a highly efficient atomistic neural network program for alloys
}
\author{Xin Chen}
\author{Xing-Yu Gao}
\author{Ya-Fan Zhao}
\author{De-Ye Lin}
\email{lindeye0716@163.com}
\author{Wei-Dong Chu}
\author{Hai-Feng Song}
\email{song\_haifeng@iapcm.ac.cn}
\affiliation{
    Institute of Applied Physics and Computational Mathematics, 
    Beijing 100088, China}
\affiliation{
    CAEP Software Center for High Performance Numerical Simulation, 
    Beijing 100088, China}

\begin{abstract}
Atomistic modeling is important for studying physical and chemical properties of
materials. Recently, machine learning interaction potentials have gained much 
more attentions as they can provide density functional theory level predictions 
within negligible time. Currently, the symmetry function descriptor based 
atomistic neural network is the most widely used model for modeling alloys. 
To precisely describe complex potential energy surfaces, integrating advanced 
metrics, such as force or virial stress, into training can be of great help. 
However, the traditional way to construct a 
machine learning model computation graph starts from pre-computed descriptors, 
but not positions, due to technical challenges. Thus, deriving analytical atomic 
force and virial stress of a machine learning model becomes quite difficult as 
the partial derivations of descriptors with respect to positions need separate 
implementations, which requires plenty of efforts and its efficiency can not be 
guaranteed.
In this paper, we designed a new method, named the virtual atom approach, for 
implementing the symmetry function descriptor and the corresponding atomistic 
neural networks. With the virtual atom approach, we are able to build 
computation graph from atomic positions \textemdash but not descriptors 
\textemdash to total energy directly, thus asking modern machine learning 
platforms to compute atomic force and virial stress becoming possible. We also 
derived a simple and machine learning friendly equation for calculating virial 
stress. All these new algorithms have been implemented in our Python package, 
TensorAlloy. TensorAlloy supports constructing machine learning interaction 
potentials for both \textit{molecules} and \textit{solids}.
\end{abstract}

@@

\newcommand{\tensoralloy}{\textbf{TensorAlloy}}

@@

In the past few years, machine learning (ML) has become one of the most 
influential topic in almost all fields of science. A lot of effort has been made 
by various physicists to explore the possiblity of using ML models \textemdash 
instead of traditional empirical potentials \textemdash to describe atomistic 
interactions during MD simulations\cite{kCON,SchNet,DTNN,SNAP_2018,SNAP_Mo_2017,
ANI,TensorMol}. ML potentials are much easier to design and 
optimize. With sufficient training data, ML models can achieve similar or even 
better performances compared with density functional theory.
The symmetry function based atomistic neural network (ANN) model, first proposed 
by Parinello and Behler in 2007\cite{Behler,Behler2,Behler3,Behler4,Behler1}, 
is still the most widely used method for \textit{solids} and its effectness and 
accuracy has been proven by several 
works\cite{Behler-solids-lzp-1,Behler-alloy-1,Behler-alloy-2, Behler-alloy-3}.
Until now, various ML implementations\cite{AMP,ANI,TensorMol} have been 
published by researchers. The Amp\cite{AMP} package, developed by Peterson's 
group, is the first open-source framework that implements the symmetry function 
descriptor. Amp can be used to learn interaction potentials for both 
\textit{molecules} and \textit{solids} and it has successfully powered quite a 
few researches since 2016\cite{Amp-works-hammer,amp-work-1}. ANI\cite{ANI} and 
TensorMol\cite{TensorMol} are also symmetry function based packages but they 
mainly focus on \textit{molecules}.

In order to precisely model atomistic interactions, both total energy and atomic 
forces are necessary. For \textit{solids}, the virial stress tensor is also an
essential metric. However, integrating force and stress into the loss function 
is not an easy task. The machine learning approaches favor vectorizable 
operations but the calculations of atomic descriptors are typically complicated 
so that many traditional ML packages (Amp, kCON, etc) choose to pre-compute 
descriptors and only build computation graph from descriptors to total energy. 
This approach can significantly reduce technical difficulty and works very well 
if only total energy is considered in the loss function. But to fit atomic 
forces, numerous codes must be written manually, such as the derivation of 
descriptors with respect to positions. Although the derivation is theoretically
clear, the implementation is challenging and the efficiency can not be 
guaranteed.

@@

@@

\newcommand{\rijn}{r_{ij\mathbf{n}}}

@@

\newcommand{\rijna}{r_{ij\mathbf{n_1}}}
\newcommand{\rikna}{r_{ik\mathbf{n_2}}}
\newcommand{\rjkna}{r_{jk\mathbf{n_3}}}

There are many choices of $F_G$. The symmetry function descriptor, proposed 
by Parinello and Behler in 2007\cite{Behler}, is currently the most widely used 
atomic descriptor, especially for alloys. The symmetry function descriptor 
consists of two sets of invariant (tranlational, rotational and permutational) 
functions: the radial symmetry function $\mathbf{G}^{(2)}$ and the angular 
symmetry function $\mathbf{G}^{(4)}$:
\begin{align}
\label{eq:G2}
\mathbf{G}^{(2)}_i(\eta, \omega) = & \sum_{j \neq i}{
    \sum_{\mathbf{n}}{g_2 (\eta, i, j, \omega)}
} \\
\label{eq:G4}
\mathbf{G}^{(4)}_i(\beta, \gamma, \zeta) 
= & 2^{1-\zeta}\sum_{j, k\neq j, k\neq i}{
    \sum_{\mathbf{n}_1}{\sum_{\mathbf{n}_2}{\sum_{\mathbf{n}_3}{
        g_4(\beta, \gamma, \zeta, 
          i, j, k, 
          \mathbf{n}_1, \mathbf{n}_2, \mathbf{n}_3)
    }}}
} \\
\label{eq:g2}
g_2(\eta, i, j, \omega) = 
& \exp\left\{-\frac{\eta(\rijn - \omega)^2}{r_c^2} \right\} \cdot f_c(\rijn) \\
\label{eq:g4}
g_4(
    \beta, \gamma, \zeta, 
    i, j, k, 
    \mathbf{n}_1, \mathbf{n}_2, \mathbf{n}_3) = & 
\left( 1 + \gamma\cos\theta \right)^\zeta 
    \exp \left\{ -\frac{\beta(\rijna^2 + \rikna^2 + \rjkna^2)}{r_c^2} \right\}
\nonumber \\
& \times f_c(\rijna)f_c(\rikna)f_c(\rjkna)
\end{align}
where $\eta$, $\omega$, $\beta$, $\gamma$ and $\zeta$ are empirically chosen 
parameters, $f_c(r)$ is a cutoff (damping) function and its original form is:
\begin{equation}
\label{eq:fc_origin}
f_c(r) =
    \begin{cases}
    0 & \text{$r > r_c$} \\
    \frac{1}{2} + \frac{1}{2}\cos\left(r / r_c \cdot \pi \right) & 
    \text{$r \leq r_c$}
    \end{cases}
\end{equation}
and $r_c$ is the cutoff radius. Equation \eqref{eq:fc_origin} can be further 
transformed to a vectorized expression:
\begin{equation}
\label{eq:fc}
f_c(r) = \frac{1}{2}\left( 
    1 + \cos\left[ \min(\frac{r}{r_c}, 1) \pi \right] 
\right)
\end{equation} 
and $\cos\theta$ also has a vectorized form:
\begin{equation}
\cos\theta = \frac{\rijna^2 + \rikna^2 - \rjkna^2}{2 \rijna \cdot \rikna}
\end{equation}
The summations in $\mathbf{G}^{(2)}$ and $\mathbf{G}^{(4)}$ should go over all 
neighbors within $r_c$.

@@

The derivations are theoretically straightforward but their implementations are 
challenging \cite{AMP}. Later we will provide a new approach to address this 
problem.

The $3 \times 3$ virial stress tensor $\epsilon$ is an important metric to 
describe solids under deformation or external pressure. For arbitrary many-body 
interaction potentials, $\epsilon$ can be calculated with the following equation
\cite{lammps_stress}:
\begin{equation}
\label{eq:stress_lammps}
V \cdot \epsilon = \left(-\sum_{i=1}^{N}{\mathbf{r}_i^{(0)} \otimes f_i} - 
\sum_{\mathbf{n}}{\mathbf{h}^T\mathbf{n}} \otimes 
\sum_{i=1}^{N}{F^{\prime}_{i\mathbf{n}}}
\right)^T
\end{equation}
where $V$ is the volume, $\otimes$ indicates tensor product, $f_i$ is the total 
force acting on atom $i$ and $F^{\prime}_{i\mathbf{n}}$ is the partial force. 
Until now, most machine learning force-field models focus on fitting total 
energy and atomic forces, very few of them\cite{DeePMD,DeePMD_rl,DeePMD_kit} had 
attempted to include stress in their loss functions. 
The current expression of Equation \ref{eq:stress_lammps} is not very compatible 
with NN or other ML potentials since the partial forces are difficult to compute 
with vectorized operations. To resolve this issue, we derived an equivalent form 
of Equation \ref{eq:stress_lammps}:
\begin{equation}
\label{eq:stress}
V \cdot \epsilon = -F^{T} R + 
\left(\frac{\partial E^{total}}{\partial \mathbf{h}}\right)^T \mathbf{h}
\end{equation}
where $R$ is the $N \times 3$ positions matrix and $F$ is the 
corresponding total forces matrix. The new equation can be easily implemented 
with simple matrix operations. The detailed derivation is given in the appendix.

@@

In this section we will describe the program design and the core algorithm: 
the virtual atom approach (VAP). The most noticable feature of the TensorAlloy 
program is that the calculation of symmetry function descriptors is finally 
implemented in the compuation graph with the help of VAP. Thus, we can utilize 
the \textbf{AutoGrad} routine of TensorFlow to derive atomic forces and virial 
stress automatically and also efficiently.

The Ni-Mo binary alloy system will be used as the example to show this 
algorithm. For simplicity, only radial symmetry function related algorithms are 
visualized in the figures. The associated codes (Python3.7) are given in the 
appendix. A demo script can be downloaded from GitHub freely 
(https://github.com/Bismarrck/vap).

Fig \ref{fig:program_design} shows the general design of TensorAlloy.
TensorAlloy has two phases: 1) is the training phase and 2) represents the 
prediction phase. These two phases have slightly different focusing points:

In the training phase, efficiency is the major concern. To accelerate model 
training, batching (or mini-batch training) is one of the most commonly used 
techniques. Batching requires aligned inputs. However, hundreds of different 
stoichiometries may exist in one dataset. To batch such structures, an universal 
approach of expressing different structures in a single reference system (the 
global symbol list or GSL) is necessary.

In the prediction phase, things may be a bit different. The stoichiometry of 
the input structure may not be included in the training dataset and it can vary 
significantly. Taking the example of Ni-Mo, the trained model must be capable of 
predicting any $\mathrm{Ni}_{x}\mathrm{Mo}_{y}$ solid while $x$ and $y$ can be 
either as large as 50000 or just zero. Hence, VAP must be flexible enough to 
handle these situations.

In the following subsection, we will discuss the details of VAP.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.A
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Preparation}
\label{section:preparation}

@@

Fig \ref{fig:Fig2} shows the initialization of VAP. The cutoff radius $r_{cut}$ 
must be fixed. The initialization involves two steps. The first step is to build 
the global symbol list (GSL), which acts as the universal reference system. A 
virtual atom is inserted at the very first position of this list. The benefits 
will be discussed later. To build GSL, all $\nelmax$ must be obtained. 
$\nelmax$ indicates the maximum appearances of an element in any structure and 
$\nel$ represents the total number of unique chemical elements:
\begin{equation}
\label{eq:n_vap}
\nvap = 1 + \sum_{\mathrm{el}}{\nelmax}
\end{equation}
For the Ni-Mo dataset, $\mathrm{el} \in [\mathrm{Ni}, \mathrm{Mo}]$ and 
$\nel = 2$. 
Table \ref{table:GSL} gives some examples of the GSL mapping. 'X' denotes the 
inserted virtual atom and Ni$_{108(54)}$ means the first 54 Ni atoms are mapped 
from the local stoichiometry. In the training phase, all $\nelmax$ are constant.
In the prediction phase, $\nelmax$ shall be determined dynamically. One should 
note that the minimum value of $\nelmax$ should be 1 due to technical reason (so 
that the computation graph of TensorFlow can be properly executed). The codes to
get $\nelmax$ and GSL are given in Appendix(C). With GSL setup, atomic positions
and forces can be transformed as well. Appendix(D) gives the implementation of 
the virtual-atom mapping mechanism.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table.1
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table}[h]
\centering
\begin{tabular}{lccccc}
\hline
 & Stoichiometry & $\nnimax$ & $\nmomax$ & $N^{\mathrm{vap}}$ 
 & Stoichiometry (GSL) \\
\hline
 & Ni$_{54}$ &  &  &  & XNi$_{108(54)}$Mo$_{176(0)}$ \\
Training & Ni$_{54}$Mo$_{12}$ & 108 & 176 & 285 
 & XNi$_{108(54)}$Mo$_{176(12)}$ \\
 & Mo$_{12}$ &  &  &  & XNi$_{108(0)}$Mo$_{176(12)}$ \\
\hline
 & Ni$_{54}$ & 54 & 1 & 56 & XNi$_{54(54)}$Mo$_{1(0)}$ \\
Prediction & Ni$_{54}$Mo$_{12}$ & 54 & 12 & 68 & XNi$_{54(54)}$Mo$_{12(12)}$ \\
 & Mo$_{12}$ & 1 & 12 & 14 & XNi$_{1(0)}$Mo$_{12(12)}$ \\
\hline
\end{tabular}
\caption{\label{table:GSL}
The mappings of local stoichiometries to their GSL forms, taking examples of the 
SNAP/Ni-Mo dataset.
}
\end{table}

Then we can determine another two key constants: $\nijmax$ and $\nijkmax$. 
Here $N_{ij}^{\mathrm{max}}$ represents the maximum number of neighbor pairs in
arbitrary structure and the subscript 'ijk' denotes triples. 
$N_{ij}^{\mathrm{max}}$ and $N_{ijk}^{\mathrm{max}}$ only depend on 
$r_{cut}$ and the associated codes are given in Appendix(B). 
Table \ref{table:nij_nijk_max} summarizes $\nijmax$ and $\nijkmax$ at different
$r_{cut}$ of the two tested datasets: QM7 and SNAP/Ni-Mo.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.B
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Interatomic distances}
\label{section:interatomic_distance}

@@

Both the radial (Equation \ref{eq:g2}) and the angular (Equation \ref{eq:g4}) 
functions only interact with interatomic distances. In order to compute symmetry
function descriptors, local atomic positions shall be transformed to GSL 
interatomic distances first. Fig \ref{fig:Fig3} demonstrates the pseudo 
computation graph for calculating interatomic distances 
$\mathbf{r}^{\mathrm{GSL}}$ for radial descriptors:

@@

    \item[2.] 
    Map $\mathbf{R}$, $\mathbf{i}$, $\mathbf{j}$ and $\mathbf{n}$ to
    their corresponding GSL forms with padding zeros. One should note that these 
    GSL arrays can be cached for training.

@@

    \item[4.]
    Compute the displacements in the original cell:
    $\mathbf{D}_{ij} = \mathbf{j}^{\mathrm{GSL}} - \mathbf{i}^{\mathrm{GSL}}$

    \item[5.]
    Compute the total periodic displacements:
    $\mathbf{D}_{pbc} = \mathbf{n}^{\mathrm{GSL}} \mathbf{h}$

@@

    \item[7.] 
    Compute the interatomic distances $\mathbf{r}^{\mathrm{GSL}}$.

@@

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.C
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Radial descriptors}
\label{section:radial_descriptors}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Fig.4
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

@@

The workflow for the batch training phase is almost identical to the prediction 
phase introduced above. There are still some noticable diiferences. Assume 
$N_{b}$ represents the batch size:

>>>>

\title{
    TensorAlloy: an automatic atomistic neural network program for alloys
}
\author{Xin Chen}
\author{Xing-Yu Gao}
\author{Ya-Fan Zhao}
\author{De-Ye Lin}
\email{lindeye0716@163.com}
\author{Wei-Dong Chu}
\author{Hai-Feng Song}
\email{song\_haifeng@iapcm.ac.cn}
\affiliation{
    Institute of Applied Physics and Computational Mathematics, 
    Beijing 100088, China}
\affiliation{
    CAEP Software Center for High Performance Numerical Simulation, 
    Beijing 100088, China}

\begin{abstract}
Atomistic modeling is important for studying physical and chemical properties of
materials. Recently, machine learning interaction potentials have gained much 
more attentions as they can provide density functional theory level predictions 
within negligible time. Currently, the symmetry function descriptor based 
atomistic neural network is the most widely used model for modeling alloys. 
To precisely describe complex potential energy surfaces, integrating advanced 
metrics, such as force or virial stress, into training can be of great help. 
The traditional way of implementing machine learning models starts from 
pre-computed descriptor and the analytical (model-derived) atomic forces and 
virial stress are handled by separate codes which require plenty of efforts. 
In this work, we choose a new route to overcome this challenge by building 
computation graph from atomic positions directly. Thus, TensorFlow or other 
modern machine learning platforms can handle the derivations of forces and 
stress, both automatically and efficiently. We demonstrate our core algorithm 
\textemdash the virtual atom approach \textemdash using Behler-Parinello's 
symmetry function atomic descriptor. This algorithm is implemented in our Python 
program TensorAlloy. TensorAlloy supports constructing machine learning 
interaction potentials for both \textit{molecules} and \textit{solids}. The QM7 
and SNAP/Ni-Mo datasets are used to discuss performances of our program.
\end{abstract}

@@

@@

In the past few years, machine learning (ML) has become one of the most 
influential topic in almost all fields of science. A lot of effort has been made 
by various physicists to explore the possiblity of using ML models \textemdash 
instead of traditional empirical potentials \textemdash to describe atomistic 
interactions during MD simulations\cite{kCON,SchNet,DTNN,SNAP_2018,SNAP_Mo_2017,
ANI,TensorMol}. ML potentials are much easier to design and 
optimize. With sufficient training data, ML models can achieve similar or even 
better performances compared with density functional theory.
The symmetry function based atomistic neural network (ANN) model, first proposed 
by Parinello and Behler in 2007\cite{Behler,Behler2,Behler3,Behler4,Behler1}, 
is still the most widely used method for \textit{solids} and its effectness and 
accuracy has been proven by several 
works\cite{Behler-solids-lzp-1,Behler-alloy-1,Behler-alloy-2, Behler-alloy-3}.
Until now, various ML implementations\cite{AMP,ANI,TensorMol} have been 
published by researchers. The Amp\cite{AMP} package, developed by Peterson's 
group, is the first open-source framework that implements the symmetry function 
descriptor. Amp can be used to learn interaction potentials for both 
\textit{molecules} and \textit{solids} and it has successfully powered quite a 
few researches since 2016\cite{Amp-works-hammer,amp-work-1}. 
TensorMol\cite{TensorMol} and its sucessor TorchANI\cite{ANI} are also symmetry 
function based ANN programs but they mainly focus on \textit{molecules}.

In order to precisely model atomistic interactions, both total energy and atomic 
forces are necessary. For \textit{solids}, the virial stress tensor is also an
essential metric. However, integrating force and stress into the loss function 
is not an easy task. The machine learning approaches favor vectorizable 
operations but the calculations of atomic descriptors are typically complicated 
so that many traditional ML packages (Amp, kCON, etc) choose to pre-compute 
descriptors and only build computation graph from descriptors to total energy. 
This traditional approach can significantly reduce technical difficulty and 
works very well if only total energy is considered in the loss function. But to 
fit atomic forces, numerous codes must be written manually, such as the 
derivation of descriptors with respect to positions. Although the derivation is 
theoreticallyclear, the implementation is challenging and the efficiency can not 
be guaranteed.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% New Commands
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\newcommand{\rijn}{r_{ij\mathbf{n}}}
\newcommand{\rijna}{r_{ij\mathbf{n_1}}}
\newcommand{\rikna}{r_{ik\mathbf{n_2}}}
\newcommand{\rjkna}{r_{jk\mathbf{n_3}}}

@@

@@

There are many choices of $F_G$. The symmetry function descriptor, proposed 
by Parinello and Behler in 2007\cite{Behler}, is currently the most widely used 
atomic descriptor. The symmetry function descriptor consists of two sets of 
invariant (tranlational, rotational and permutational) functions: the radial 
symmetry function $\mathbf{G}^{(2)}$ and the angular symmetry function 
$\mathbf{G}^{(4)}$:
\begin{align}
\label{eq:G2}
\mathbf{G}^{(2)}_i(\eta, \omega) = & \sum_{j \neq i}{
    \sum_{\mathbf{n}}{g_2 (\eta, i, j, \omega)}
} \\
\label{eq:G4}
\mathbf{G}^{(4)}_i(\beta, \gamma, \zeta) 
= & 2^{1-\zeta}\sum_{j, k\neq j, k\neq i}{
    \sum_{\mathbf{n}_1}{\sum_{\mathbf{n}_2}{\sum_{\mathbf{n}_3}{
        g_4(\beta, \gamma, \zeta, 
          i, j, k, 
          \mathbf{n}_1, \mathbf{n}_2, \mathbf{n}_3)
    }}}
} \\
\label{eq:g2}
g_2(\eta, i, j, \omega) = 
& \exp\left\{-\frac{\eta(\rijn - \omega)^2}{r_c^2} \right\} \cdot f_c(\rijn) \\
\label{eq:g4}
g_4(
    \beta, \gamma, \zeta, 
    i, j, k, 
    \mathbf{n}_1, \mathbf{n}_2, \mathbf{n}_3) = & 
\left( 1 + \gamma\cos\theta \right)^\zeta 
    \exp \left\{ -\frac{\beta(\rijna^2 + \rikna^2 + \rjkna^2)}{r_c^2} \right\}
\nonumber \\
& \times f_c(\rijna)f_c(\rikna)f_c(\rjkna)
\end{align}
where $\eta$, $\omega$, $\beta$, $\gamma$ and $\zeta$ are empirically chosen 
parameters, $f_c(r)$ is a cutoff (damping) function and its original form is:
\begin{equation}
\label{eq:fc_origin}
f_c(r) =
    \begin{cases}
    0 & \text{$r > r_c$} \\
    \frac{1}{2} + \frac{1}{2}\cos\left(r / r_c \cdot \pi \right) & 
    \text{$r \leq r_c$}
    \end{cases}
\end{equation}
and $r_c$ is the cutoff radius. Equation \eqref{eq:fc_origin} can be further 
transformed to a vectorized expression:
\begin{equation}
\label{eq:fc}
f_c(r) = \frac{1}{2}\left( 
    1 + \cos\left[ \min(\frac{r}{r_c}, 1) \pi \right] 
\right)
\end{equation} 
and $\cos\theta$ also has a vectorized form:
\begin{equation}
\cos\theta = \frac{\rijna^2 + \rikna^2 - \rjkna^2}{2 \rijna \cdot \rikna}
\end{equation}
The summations in $\mathbf{G}^{(2)}$ and $\mathbf{G}^{(4)}$ should go over all 
neighbors within $r_c$.

@@

The derivations are theoretically straightforward but their implementations are 
challenging \cite{AMP}. Later we will provide an automatic approach to address 
this problem.

The $3 \times 3$ virial stress tensor $\epsilon$ is an important metric to 
describe solids under deformation or external pressure. For arbitrary many-body 
interaction potentials, $\epsilon$ can be calculated with the following equation
\cite{lammps_stress}:
\begin{equation}
\label{eq:stress_lammps}
V \cdot \epsilon = \left(-\sum_{i=1}^{N}{\mathbf{r}_i^{(0)} \otimes f_i} - 
\sum_{\mathbf{n}}{\mathbf{h}^T\mathbf{n}} \otimes 
\sum_{i=1}^{N}{F^{\prime}_{i\mathbf{n}}}
\right)^T
\end{equation}
where $V$ is the volume, $\otimes$ indicates tensor product, $f_i$ is the total 
force acting on atom $i$ and $F^{\prime}_{i\mathbf{n}}$ is the partial force. 
Until now, most machine learning force-field models focus on fitting total 
energy and atomic forces, very few of them\cite{DeePMD,DeePMD_rl,DeePMD_kit} had 
attempted to include stress in their loss functions. 
The current expression of Equation \ref{eq:stress_lammps} is not very compatible 
with NN or other ML potentials since the partial forces are difficult to compute 
with vectorized operations. However, we find an "equivalent" form 
of Equation \ref{eq:stress_lammps}:
\begin{equation}
\label{eq:stress}
V \cdot \epsilon = -F^{T} R + 
\left(\frac{\partial E^{total}}{\partial \mathbf{h}}\right)^T \mathbf{h}
\end{equation}
where $R$ is the $N \times 3$ positions matrix and $F$ is the 
corresponding total forces matrix. This new form can be easily implemented 
within TensorFlow. The detailed derivation is given in the appendix.

@@

In this section we will describe the architecture of TensorAlloy and the core 
algorithm: the virtual atom approach (VAP). The most noticable feature of the 
TensorAlloy program is that the calculation of symmetry function descriptors is 
automatically handled by the \textbf{AutoGrad}\cite{tensorflow} routine of 
TensorFlow because we successfully find a route (VAP) to build the direct 
computation graph from atomic positions to total energy.

The Ni-Mo binary alloy system will be used as the example to show our 
algorithms. The original dataset (SNAP) is published by Shyue Ping 
Ong\cite{SNAP_2018}. 
For simplicity, only radial symmetry function related algorithms are 
visualized in the figures. Python3.7 implementations of key algorithms and 
functions are given in the appendix. A full demo (including the alloy datasets) 
can be obtained from GitHub freely (https://github.com/Bismarrck/vap).

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.A
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{General design}
\label{section:principles}

Fig \ref{fig:program_design} shows the general design of TensorAlloy.
TensorAlloy has two phases: the training phase \textbf{(1)} and the prediction 
phase \textbf{(2)}. These two phases share many common requirements, but they 
also have their own concerns.

For the training phase, in order to make TensorAlloy a general-purpose and
efficient machine learning potential training program, these three key 
challenges shall be resolved:
\begin{itemize}

    \item[1.] 
    Stoichiometry-free: the training dataset should not have any stoichiometry 
    restriction. Any type of \textit{solid} or \textit{molecule} is acceptable. 
    A universal approach of expressing different structures in a single 
    reference system may be necessary.

    \item[2.] 
    Mini-batch training: mini-batch based stochastic training is 
    currently the most efficient way to train neural networks on large datasets.
    However, batch training requires \textit{vectorized} and \textit{aligned} 
    expressions. Here \textit{aligned} means feature arrays of structures of 
    different stoichiometries share the same shape.

    \item[3.] 
    Cache: some intermediate arrays may be pre-computed and stored in cache 
    files. During training, these values can be loaded from cache directly, thus 
    saving significantly amount of resources.

In the prediction phase, the situation is a bit different. The stoichiometry of 
the target structure or molecule may not be included in the training dataset and 
it can vary significantly. TensorAlloy must be capable of handling such 
(extreme) situations. For example, the trained Ni-Mo shall be applicable to any 
any $\mathrm{Ni}_{x}\mathrm{Mo}_{y}$ structure while $x$ and $y$ can be 
either as large as 100000 or as small as 0.

Besides, we also want our TensorAlloy program have an \textit{independent} and
\textit{user-friendly} prediction phase. Users can just focus on the integration 
of machine learning potentials with their own researches.

The virtual-atom approach plays a central role in handling all the challenges 
above and unify the training and prediction phases. In the following sections, 
we will gradually discuss the details of this approach.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.B
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Preparation}
\label{section:preparation}

@@

Fig \ref{fig:Fig2} shows the preparation stage of VAP. The cutoff radius 
$r_{cut}$ must be fixed. This stage has two major steps. The first step is to 
build the global symbol list (GSL), which acts as the universal reference 
system. A virtual atom is inserted at the very first position of GSL. The 
benefits will be discussed later. To build GSL, all $\{\nelmax\}$ must be 
pre-determined where $\nelmax$ is the maximum appearances of element 
\textit{el} in any structure. $\nel$ represents the number of unique chemical 
elements in the dataset. Then, $\nvap$, the dimension of GSL, can be determined:
\begin{equation}
\label{eq:n_vap}
\nvap = 1 + \sum_{\mathrm{el}}{\nelmax}
\end{equation}
For the Ni-Mo dataset, $el \in [\mathrm{Ni}, \mathrm{Mo}]$ and 
$\nel = 2$. 
Table \ref{table:GSL} gives some examples of the GSL mapping. 'X' denotes the 
inserted virtual atom and Ni$_{108(54)}$ means the first 54 Ni atoms are mapped 
from the local stoichiometry. In the training phase, all $\nelmax$ are fixed.
In the prediction phase, $\nelmax$ shall be determined dynamically. One should 
note that the minimum value of $\nelmax$ should be 1 due to technical reason (so 
that the computation graph of TensorFlow can be properly executed). The codes to
get $\nelmax$ and GSL are given in Appendix(C). With GSL setup, atomic positions
and forces can be transformed as well. Appendix(D) gives the implementation of 
the virtual-atom mapping mechanism.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table.1
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table}[h]
\centering
\begin{tabular}{lccccc}
\hline
 & Stoichiometry (Local) & $\nnimax$ & $\nmomax$ & $N^{\mathrm{vap}}$ 
 & Stoichiometry (GSL) \\
\hline
 & Ni$_{54}$ &  &  &  & XNi$_{108(54)}$Mo$_{176(0)}$ \\
Training & Ni$_{54}$Mo$_{12}$ & 108 & 176 & 285 
 & XNi$_{108(54)}$Mo$_{176(12)}$ \\
 & Mo$_{12}$ &  &  &  & XNi$_{108(0)}$Mo$_{176(12)}$ \\
\hline
 & Ni$_{54}$ & 54 & 1 & 56 & XNi$_{54(54)}$Mo$_{1(0)}$ \\
Prediction & Ni$_{54}$Mo$_{12}$ & 54 & 12 & 68 & XNi$_{54(54)}$Mo$_{12(12)}$ \\
 & Mo$_{12}$ & 1 & 12 & 14 & XNi$_{1(0)}$Mo$_{12(12)}$ \\
\hline
\end{tabular}
\caption{\label{table:GSL}
The mappings of local stoichiometries to their GSL forms, taking examples of the 
SNAP/Ni-Mo dataset.
}
\end{table}

Then we can determine another two key constants: $\nijmax$ and $\nijkmax$. 
Here $\nijmax$ represents the maximum number of atom-atom pairs in
arbitrary structure and the subscript 'ijk' denotes triples 
(or three-atoms interactions). $\nijmax$ and $\nijkmax$ only depend on 
$r_{cut}$ and the associated codes are given in Appendix(B). 
Table \ref{table:nij_nijk_max} summarizes $\nijmax$ and $\nijkmax$ at different
$r_{cut}$ of the two demo datasets: QM7 and Ni-Mo (SNAP).

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.C
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Interatomic distances}
\label{section:interatomic_distance}

@@

Both the radial (Equation \ref{eq:g2}) and the angular (Equation \ref{eq:g4}) 
symmetry functions interact with interatomic distances only. In order to compute 
the descriptors, local atomic positions should be transformed to interatomic 
distances first. Fig \ref{fig:Fig3} demonstrates the pseudo graph of 
$\mathbf{R} \rightarrow \mathbf{r}^{\mathrm{GSL}}$ for radial symmetry function
descriptors:

@@

    \item[2.] 
    Transform and expand local $\mathbf{R}$, $\mathbf{i}$, $\mathbf{j}$ and 
    $\mathbf{n}$ to their corresponding GSL arrays. These GSL arrays can be 
    cached for training. The transformations of $\mathbf{i}$, $\mathbf{j}$ and 
    $\mathbf{n}$ are straightforward: just appending $\nijmax - N_{ij}$ trailing 
    zeros. Fig \ref{fig:Fig3si} shows the mapping of 
    $\mathbf{R} \rightarrow \mathbf{R}^{\mathrm{GSL}}$ taking the example of 
    Ni$_3$Mo$_2$.

@@

    \item[4.]
    Compute the displacement vectors in the original cell:
    $\mathbf{D}_{ij} = \mathbf{j}^{\mathrm{GSL}} - \mathbf{i}^{\mathrm{GSL}}$

    \item[5.]
    Compute the periodic displacements:
    $\mathbf{D}_{pbc} = \mathbf{n}^{\mathrm{GSL}} \mathbf{h}$

@@

    \item[7.] 
    Compute the interatomic distances $\mathbf{r}^{\mathrm{GSL}}$ by calculating
    the 2-norm of each row of matrix $\mathbf{D}^{\mathrm{GSL}}$.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Fig.3 (SI)
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[h!]
\centering
\includegraphics[scale=0.8]{figures/Fig3si.pdf}
\caption{\label{fig:Fig3si} The transformation of local $\mathbf{R}$ of 
Ni$_3$Mo$_2$ to $\mathbf{R}^{\mathrm{GSL}}$ of XNi$_{108(3)}$Mo$_{176(2)}$.
}
\end{figure}

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.D
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Radial descriptors}
\label{section:radial_descriptors}

@@

The workflow for the batch training phase is almost identical to the prediction 
phase described above. Meanwhile, there are some noticable diiferences:
